Comparison among doublet predictions on benchmarking datasets

### 2024-06-04 analyses/doublet-detection/exploratory-notebooks/03_compare-benchmark-results.Rmd
Some reorg for logical flow, intro and conclusions section

@@

### Read and prepare input data

First, we'll read in and combine doublet results into a list of data frames for each dataset.
We'll also two columns for each dataset:

@@

## Performance metrics

This section shows performance metrics for the consensus calls for each dataset.

```{r}
doublet_df_list |>
  purrr::iwalk( 
    \(df, dataset) {
        print(glue::glue("======================== {dataset} ========================"))

        caret::confusionMatrix(
          # truth should be first
          table(
            "Truth" = df$ground_truth,
            "Consensus prediction" = df$consensus_call
          ), 
          positive = "doublet"
        ) |>
        print()
    }
  )
```

## Visualizations

@@

This section plots the PCA for each dataset, with three color schemes from left to right:
- Ground truth doublets are shown in black
- Consensus doublets are shown in black
- Points are colored as tp, tn, fp, or fn based on comparing the consensus call to the ground truth

@@

### Upset plots

This section shows upset plots for overlap among methods for each dataset.

>>>>

This notebook compares doublet results calculated on benchmarking datasets to one another, with the primary goal of addressing these questions:

1. Do methods tend to predict overlapping or distinct sets of doublets on the same dataset?
2. Is the consensus doublet call across methods predictive of the true doublet status?
Here, the "consensus doublets" are those droplets which all methods identify as doublets.

There are several items to bear in mind when interpreting these results:

- The ground truth calls themselves to which we are comparing consensus calls may not be entirely accurate, since they were also computationally identified generally with demultiplexing algorithms. 
- `cxds`, as we are using it, does not have a specific threshold for calling droplets. 
By contrast, both `scublet` and `scDblFinder` identify a threshold based on the given dataset they are processing. 
This notebook used a doublet threshold of `>=0.75` for `cxds`, which may not be universally suitable (though choosing a universally suitable threshold is not easy either!).

@@

## Read and prepare input data

First, we'll read in and combine doublet results into a list of data frames for each dataset.
We'll also create new columns for each dataset:

@@

## Upset plots

This section contains upset plots that show overlap across doublet calls from each method, displayed for each dataset.

## Evaluating consensus performance

This section visualizes and evaluates the consensus doublet calls across each dataset.

@@

This section plots the PCA for each dataset, with three color schemes from left to right:

1. Ground truth doublets are shown in black and singlets in blue
2. Consensus doublets are shown in black and all remaining droplets in blue
3. Points are colored based on comparing the consensus call to the ground truth as one of:
  - true positive (`tp`), true negative (`tn`), false positive (`fp`), false negative (`fn`)

@@

### Performance metrics

This section calculates a confusion matrix and associated statistics on the consensus calls.

```{r}
metric_df <- doublet_df_list |>
  purrr::imap( 
    \(df, dataset) {
        print(glue::glue("======================== {dataset} ========================"))

        confusion_result <- caret::confusionMatrix(
          # truth should be first
          table(
            "Truth" = df$ground_truth,
            "Consensus prediction" = df$consensus_call
          ), 
          positive = "doublet"
        )

        print(confusion_result)

        # Extract information we want to present later in a table
        tibble::tibble(
          "Dataset name" = dataset,
          "Kappa" = round(confusion_result$overall["Kappa"], 3), 
          "Balanced accuracy" = round(confusion_result$byClass["Balanced Accuracy"], 3)
        )
    }
  ) |>
  dplyr::bind_rows()
```

## Conclusions

Overall, methods do not have substantial overlap with out another. 
They each tend to detect different sets of doublets, leading to fairly small sets of consensus doublets. 
Further, the consensus doublets called by all three methods have some, but not substantial, overlap with the ground truth.

For three out of four datasets, `scDblFinder` predicts a much larger number of doublets compared to other methods. 
For `pdx-MULTI`, however, `cxds` predicts a much larger number of droplets.

The table below summarizes performance of the "consensus caller". 
Note that, in the [benchmarking paper these datasets were originally analyzed in](https://doi.org/10.1016/j.cels.2020.11.008), `hm-6k` was observed to be one of the "easiest" datasets to classify across methods.  
Consistent with that observation, it has the highest `kappa` value here, although it is still fairly low - though not as low as the other methods, which are very close to 0.

```{r}
metric_df
```
